---
title: "assign2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
library(tidyverse)
candyBars <- read_csv('C:/Users/jonathan/Desktop/DAPT Docs/spring 2019/Statistics/stat_assignments_spring19_R/stat_assignment1_R/data/CandyBars.txt')
library(corrplot)
library(ggplot2)
library(ggbiplot)
library(kableExtra)
library(nFactors)
library(FactoMineR)
```

# Part 1

Perform a Factor Analysis on this data by carrying out the following steps:
1) How many factors should be retained (based on eigenvalues and a Scree plot)? You may use
any “Factoring method” or “Rotation method” you wish. The defaults, of course, are fine.
2) Provide the rotated factor loadings matrix. You should notice that (as compared with Principal
Component Analysis) fewer variables load onto each factor. Interpret (as best as possible) your
factors. Can you give an overall name to each factor?
3) Plot factor scores for your first two factors. Color the points by Brand. Are there any natural
groupings of observations?

```{r factor}
#subset numeric variables
candyBars_sub <- candyBars %>% 
  select_if(is.numeric)

ev_candyBars <- eigen(cor(candyBars_sub))

ap <- parallel(subject=nrow(candyBars_sub),var=ncol(candyBars_sub),
  rep=100,cent=.05)
nS <- nScree(x=ev_candyBars$values, aparallel=ap$eigen$qevpea)
plotnScree(nS)

candyfact <- factanal(candyBars_sub, factors = 5)

factorplot <- PCA(candyBars_sub)

candyfact$loadings

```

# Part 2

1) Perform a linear discriminant analysis using the highlighted variables for discriminating the
variable Pop_Size_Group.
2) Show a plot of the two discriminant functions with points colored by Pop_Size_Group. Do
both discriminant functions appear necessary to describe group separation?
3) Give the scoring coefficients for the two discriminant functions. Which variables
appear most important for describing group separation?
4) Provide a classification matrix and the percent of observations misclassified. 

```{r disc}
library(MASS)
demographics <- read_csv('C:/Users/jonathan/Desktop/DAPT Docs/spring 2019/Statistics/stat_assignments_spring19_R/stat_assignment2_R/data/Demographics.txt')
demdisc <- lda(Pop_Size_Group ~ ., data=demographics, na.action = "na.omit")

#Percent Correct
demdisc.values <- predict(demdisc,demographics)
ct <-table(demographics$Pop_Size_Group,demdisc.values$class)
sum(diag(prop.table(ct)))

plot(demdisc)

```